sWGS QC

### 2020-04-21 Paper.Rmd
Updating figures

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(BarrettsProgressionRisk)
library(ggrepel)
library(kableExtra)
library(reshape2)
library(gridExtra)
library(RColorBrewer)
library(ggfortify)

@@

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T)

pROC::coords(rocNOHGD, 'best', transpose=F) %>% add_column('AUC'=pROC::auc(rocNOHGD)) %>% mutate_all(list(~round(.,2))) %>% 
  kable(caption = 'LOO excluding HGD/IMC') %>% kable_styling(full_width = F)

ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T) , width=6, height=6, units='in', dpi=300)
```

@@

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonolgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

@@

rocNOLGD = pROC::roc(Status ~ Prediction, data=pg.sampNOLGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',incS = T)

pROC::coords(rocNOLGD, 'best', transpose=F) %>% add_column('AUC'=pROC::auc(rocNOLGD)) %>% mutate_all(list(~round(.,2))) %>% 
  kable(caption = 'LOO excluding LGD/HGD/IMC') %>% kable_styling(full_width = F)

ggsave(filename='plots/auc/roc_nolgd_model.png', plot=roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',incS = T) , width=6, height=6, units='in', dpi=300)
```

m = bind_rows(
      get.roc.stat(roc),
      get.roc.stat(rocNOHGD)
#      get.roc.stat(rocNOLGD)
      ) %>% dplyr::mutate(model = factor(model)) %>% dplyr::mutate_if(is.numeric, list(~round(.,2)))

p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
  geom_errorbar(position=dodge, width=0.3, color='black' ) +
  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer(palette='Set2') +
  theme(legend.position = 'right',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
  labs(x='', y='ROC', title='Model comparisons')

auc.leg = get.legend(p)

ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=6, width=4.5, units="in", limitsize=F, dpi = 300)

@@

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
rocEndoM$model = 'Max per Endoscopy'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
rocEndoA$model = 'Mean per Endoscopy'
ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Mean per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
rocLastEndo$model = 'Last endoscopy'
ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

pendo.rand = pg.samp %>% dplyr::group_by(Patient, Status, PID) %>%
  dplyr::summarise( r.Prob = sample(Prediction, 1), n = length(PID) ) %>% filter(n > 1)
rocRandEndo = pROC::roc(Status ~ r.Prob, data=pendo.rand, ci=T, of='thresholds')
rocRandEndo$model = 'Per Endoscopy Random'
ggsave('plots/auc/randomendo_auc.png', plot=roc.plot(rocRandEndo, title='Random per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

endo.thresholds = bind_cols('roc'=c('Max','Mean','Last','First','Random'),
  bind_rows(
    pROC::coords(rocEndoM, 'best', transpose = T),
    pROC::coords(rocEndoA, 'best', transpose = T),
    pROC::coords(rocLastEndo, 'best', transpose = T),
    pROC::coords(rocFirstEndo, 'best', transpose = T),
    pROC::coords(rocRandEndo, 'best', transpose = T)
  ))

multi.roc.plot(list(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo)) + labs(title='Endoscopy prediction ROCs')
```

@@

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
#pROC::coords(rocPtM, 'best', transpose=T)
rocPtM$model = 'Max per Patient'
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
#pROC::coords(rocPtA, 'best', transpose=T))
rocPtA$model = 'Mean per Patient'
ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Mean per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

grid.arrange(roc.plot(rocPtM, title='Max') , roc.plot(rocPtA, title='Mean'), nrow=1, top='Per Patient')

multi.roc.plot(list(rocPtM, rocPtA)) + labs(title='Per-Patient ROCs')

@@

roc$model = "CN Model"
m = bind_rows(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc)) %>%
  dplyr::mutate(model = factor(model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)) %>% 
  dplyr::mutate_if(is.numeric, list(~round(.,2)))

rocList = list(roc,pathroc,hgdpathroc,p53roc)

@@

@@

m = bind_rows(
      get.roc.stat(roc),
      get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
      get.roc.stat(rocPtM), get.roc.stat(rocPtA)
      #get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo),
      #get.roc.stat(rocRandEndo)
      ) %>% 
  dplyr::mutate_if(is.numeric, list(~round(.,2))) %>%
  dplyr::mutate(model = factor(model, levels=c('per Sample','Mean per Endoscopy','Max per Endoscopy','Mean per Patient','Max per Patient')))

p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
  geom_errorbar(position=dodge, width=0.3, color='black' ) +
  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer(palette='Set2') +
  theme(legend.position = 'bottom',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
  labs(x='', y='AUC', title='Aggregated ROC Comparison')

@@

@@

ogjLimits = levels(factor(pg.samp$ogj, ordered = T))

# for (i in sort(unique(sum.patient.data$Patient))) {
#   print(i)
#   pt = filter(sum.patient.data, Patient == i)
#   status = as.character(filter(sum.patient.data, Patient == i)$Status)
# 
#   pms = mtn.spatial(filter(globalCNdf, Patient == i & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
# 
#   fl = paste0('patient_', i, '_',status,'.png')
#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
# }
p2 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 20 (P)' )
p6 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
ggsave(filename='plots/spatial/patient_20_P.png', plot=grid.arrange(p2$grob), width=10.5, height=4.5, units='in', dpi=300)
ggsave(filename='plots/spatial/patient_51_NP.png', plot=grid.arrange(p6$grob), width=10.5, height=4.5, units='in', dpi=300)

grid.arrange(p2$grob, p6$grob, ncol=1)

@@

ggsave('plots/spatial/legend.png', plot=grid.arrange(p2$legend), width=2,height=2, units='in', dpi=300)
```

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
  geom_errorbar(color='grey55', width=0.2) +
  geom_text(aes(label=paste(round(pred.r*100), '%', sep='')), color='black', nudge_y=2) +
  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
  labs(x='Months prior to endpoint', y='Percent classified', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)

@@

val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% tally %>% spread(Status, n)
val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Sex) %>% tally %>% spread(Status, n)

val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Smoking) %>% tally %>% spread(Status, n)

@@

create.desc<-function(pt,age,sex,smk,len,date) {
  str =  paste0('Patient ', pt, ': ', age, sex)
  if (smk != '' & !is.na(smk)) str = paste0(str, ' ,', smk)
  str = paste0(str,'\n')

@@

pt = 34
p34 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34, width=6.35, height=4.5,dpi=72)

pt = 3
p3 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p3, width=6.35, height=4,dpi=72)

pt = 68
p68 = vig.plot(df=filter(pg.samp, Patient == pt),info= filter(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p68, width=6.35, height=5,dpi=72)

@@

p = second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% dplyr::rename(RFA = `1`, `6-12 month recheck` = `2`, `12-24 month recheck` = `3`, `3-5 year surveillance` = `4`) %>% melt(id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
  ggplot(aes(variable, value, group = Status)) + 
  geom_col(aes(fill=Status), position = position_dodge()) + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1)) + coord_flip() +
  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + labs(x='', y='# Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
p

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(broom)
library(BarrettsProgressionRisk)
library(ggrepel)
library(kableExtra)
library(reshape2)
library(gridExtra)
library(RColorBrewer)
library(ggfortify)

@@

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD, 'Model excluding HGD/IMC',incS = T, 0.3)

pROC::coords(rocNOHGD, 0.3, transpose=F) %>% add_column('AUC'=pROC::auc(rocNOHGD)) %>% mutate_all(list(~round(.,2))) %>% 
  kable(caption = 'LOO excluding HGD/IMC') %>% kable_styling(full_width = F)

ggsave(filename='plots/auc/roc_nohgd_model.png', plot=roc.plot(rocNOHGD, 'Model excluding HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)
```

@@

```{r loonolgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonolgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

@@

rocNOLGD = pROC::roc(Status ~ Prediction, data=pg.sampNOLGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',T,0.3)

pROC::coords(rocNOLGD, 0.3, transpose=F) %>% add_column('AUC'=pROC::auc(rocNOLGD)) %>% mutate_all(list(~round(.,2))) %>% 
  kable(caption = 'LOO excluding LGD/HGD/IMC') %>% kable_styling(full_width = F)

ggsave(filename='plots/auc/roc_nolgd_model.png', plot=roc.plot(rocNOLGD, 'Model excluding LGD/HGD/IMC',T, 0.3) , width=6, height=6, units='in', dpi=300)
```

@@

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
rocEndoM$model = 'Max per Endoscopy'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
rocEndoA$model = 'Mean per Endoscopy'
ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Mean per Endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
rocLastEndo$model = 'Last endoscopy'
ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)

pendo.rand = pg.samp %>% dplyr::group_by(Patient, Status, PID) %>%
  dplyr::summarise( r.Prob = sample(Prediction, 1), n = length(PID) ) %>% filter(n > 1)
rocRandEndo = pROC::roc(Status ~ r.Prob, data=pendo.rand, ci=T, of='thresholds')
rocRandEndo$model = 'Per Endoscopy Random'
ggsave('plots/auc/randomendo_auc.png', plot=roc.plot(rocRandEndo, title='Random per Endoscopy', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)

# endo.thresholds = bind_cols('roc'=c('Max','Mean','Last','First','Random'),
#   bind_rows(
#     pROC::coords(rocEndoM, 'best', transpose = T),
#     pROC::coords(rocEndoA, 'best', transpose = T),
#     pROC::coords(rocLastEndo, 'best', transpose = T),
#     pROC::coords(rocFirstEndo, 'best', transpose = T),
#     pROC::coords(rocRandEndo, 'best', transpose = T)
#   ))

multi.roc.plot(list(rocEndoM, rocEndoA, rocLastEndo, rocFirstEndo, rocRandEndo), threshold = 0.3) + labs(title='Endoscopy prediction ROCs')
```

@@

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
#pROC::coords(rocPtM, 'best', transpose=T)
rocPtM$model = 'Max per Patient'
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
#pROC::coords(rocPtA, 'best', transpose=T))
rocPtA$model = 'Mean per Patient'
ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Mean per Patient (excl. final)', T, 0.3), width=6, height=6, units="in", limitsize=F, dpi = 300)

multi.roc.plot(list(rocPtM, rocPtA), threshold = 0.3) + labs(title='Per-Patient ROCs', subtitle = 'threshold = 0.3')
```

@@

roc$model = "CN Model"
m = bind_rows(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc, 0.3)) %>%
  dplyr::mutate(model = factor(model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)) %>% 
  dplyr::mutate_if(is.numeric, list(~round(.,2)))

@@

m = bind_rows(
      get.roc.stat(roc,0.3),
      get.roc.stat(rocNOHGD,0.3)
#      get.roc.stat(rocNOLGD)
      ) %>% dplyr::mutate(model = factor(model)) %>% dplyr::mutate_if(is.numeric, list(~round(.,2)))

p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
  geom_errorbar(position=dodge, width=0.3, color='black' ) +
  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer(palette='Set2') +
  theme(legend.position = 'right',text = element_text(size=12)) + 
  labs(x='', y='ROC', title='LOO model comparisons', subtitle = 'threshold = 0.3') 
auc.leg = get.legend(p)

#p = multi.roc.plot(list(roc,rocNOHGD), threshold = 0.3, palette = 'Dark2' ) + labs(title='LOO model comparisons', subtitle = 'threshold = 0.3')
p

#ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=4, width=4, units="in", limitsize=F, dpi = 300)

ggsave('plots/auc/loo_model_roc.png', plot=p + theme(legend.position = 'none'), height=4, width=4, units="in", limitsize=F, dpi = 300)

@@

multi.roc.plot(list(roc,rocEndoM, rocEndoA), threshold = 0.3, palette = 'Dark2' ) + labs(title='Aggregated per Endoscopy ', subtitle = 'threshold = 0.3')

m = bind_rows(
      get.roc.stat(roc,0.3),
      get.roc.stat(rocEndoM,0.3),get.roc.stat(rocEndoA,0.3),
      get.roc.stat(rocPtM,0.3), get.roc.stat(rocPtA,0.3)
      #get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo),
      #get.roc.stat(rocRandEndo)
      ) %>% 
  dplyr::mutate_if(is.numeric, list(~round(.,2))) %>%
  dplyr::mutate(model = factor(model, levels=c('per Sample','Mean per Endoscopy','Max per Endoscopy','Mean per Patient','Max per Patient')))

p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
  geom_errorbar(position=dodge, width=0.3, color='black' ) +
  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer(palette='Set2') +
  theme(legend.position = 'bottom',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
  labs(x='', y='ROC ', title='Aggregated ROC Comparison', subtitle='threshold = 0.3')

@@

p

@@

ogjLimits = levels(factor(pg.samp$ogj, ordered = T))
```

```{r}
# for (i in sort(unique(sum.patient.data$Patient))) {
#   print(i)
#   pt = filter(sum.patient.data, Patient == i)
#   status = as.character(filter(sum.patient.data, Patient == i)$Status)
# 
#   pms = mtn.spatial(filter(globalCNdf, Patient == i & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
# 
#   fl = paste0('patient_', i, '_',status,'.png')
#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
# }
p20 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits )
ggsave(filename='plots/spatial/patient_20_P.png', plot=grid.arrange(p20$grob), width=16, height=6, units='in', dpi=300)

p51 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
ggsave(filename='plots/spatial/patient_51_NP.png', plot=grid.arrange(p51$grob), width=10.5, height=4.5, units='in', dpi=300)

grid.arrange(p20$grob, p51$grob, ncol=1)

@@

ggsave('plots/spatial/legend.png', plot=grid.arrange(p20$legend), width=2,height=2, units='in', dpi=300)
```

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
  geom_errorbar(color='grey55', width=0.2) +
  geom_text(aes(label=paste(round(pred.r*100), '%', sep='')), color='black', nudge_y=2) +
  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
  labs(x='Months prior to endpoint', y='Sensitivity', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)

@@

val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status) %>% tally %>% spread(Status, n)
val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Sex) %>% tally %>% spread(Status, n)
chisq.test(table(val.demo$Sex, val.demo$Status))

val.demo %>% filter(Patient != no_followup$Patient) %>% group_by(Status,Smoking) %>% tally %>% spread(Status, n) 
chisq.test(table(val.demo$Smoking, val.demo$Status)[-3,])

@@

create.desc<-function(pt,age,sex,smk,len,date) {
  str =  paste0('Patient ', pt, ': ', age, sex)
  if (smk != '' & !is.na(smk)) str = paste0(str, ', ', smk)
  str = paste0(str,'\n')

@@

pt = 34

p34 = vig.plot(df=filter(pg.samp, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom', text = element_text(size=12)) + labs(title = filter(sum.patient.data, Patient == pt)$desc)

#vig.legend = get.legend(p34)
#ggsave('plots/vignette/legend_long.png', plot=grid.arrange(vig.legend), width=8, height=1)
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34 + theme(legend.position = 'none'), width=6.35, height=4.5,dpi=300)

pt = 3
p3 = vig.plot(df=filter(pg.samp, Patient == pt),cols=riskCols) + 
    labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p3, width=6.35, height=4,dpi=300)

pt = 68
p68 = vig.plot(df=filter(pg.samp, Patient == pt),cols=riskCols) + 
    labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p68, width=6.35, height=6,dpi=300)

@@

p = second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% dplyr::rename(RFA = `1`, `6-12 month recheck` = `2`, `12-24 month recheck` = `3`, `3-5 year surveillance` = `4`) %>% melt(id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
  ggplot(aes(variable, value, group = Status)) + 
  geom_col(aes(fill=Status), position = position_dodge()) + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1)) + coord_flip() +
  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + labs(x='', y='# Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
p
ggsave('plots/second_endo_rx.png', plot=p, width=6, height=6, units='in', dpi=300)

### 2020-04-21 sWGS QC.Rmd
Updating figures

```{r, echo=F, fig.height=10}
res.variance = arrange(res.variance,varMAD_median)
cutoff = median(res.variance$varMAD_median)+sd(res.variance$varMAD_median)*2
cutoff = 0.008

@@

## Validation dataset

@@

all.val = all.val %>% rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup
all.val = all.val %>% mutate(
  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
  Sample = paste(`SLX-ID`,sub('-','_',`Index Sequence`),sep='.')
  )

resVal = do.call(bind_rows, lapply(list.files(dir2, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types = 'ccddddl') #%>% set_names('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3', 'n.segs', 'Pass') %>%  
   # mutate(`Patient ID` = as.character(basename(dirname(f))))
})) %>% dplyr::select(patient, samplename, matches('varMad'), Pass)

resVal = resVal %>% filter(samplename %in% all.val$Sample)

resVal %>% kable(caption='Residuals across segments per sample') %>% kable_styling('striped')
resVal = arrange(resVal,varMAD_median)

ggplot(resVal,aes(sample=varMAD_median)) + stat_qq() + geom_hline(yintercept=cutoff, linetype='dashed',color='grey') + 
  labs(title='Normal Q-Q plot (all)', x='Theoretical quantiles',y='Sample quantiles') + theme_bw()

ggplot(resVal, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
  geom_vline(xintercept = cutoff, linetype='dashed', color='grey') + labs(x='var(median(segments)+1sd(segments))') +
  theme_bw()

failedQCVal = resVal %>% filter(varMAD_median > cutoff) %>% arrange(`Patient ID`, varMAD_median)

@@

The result from this is that `r nrow(failedQCVal)` samples (`r signif(nrow(failedQCVal)/nrow(resVal)*100,3)`%) would have failed due to high variance across all the segments. A manual check of these agrees with all of them.

## If we learned it from all residuals

```{r}
res.variance = res.variance %>% mutate_at(vars(patient, sample), as.character)
resVal = resVal %>% dplyr::rename( patient = `Patient ID` ) %>% mutate_at(vars(patient, sample), as.character)

res.variance$cohort = 'discovery'
resVal$cohort = 'validation'

ar = bind_rows( res.variance, resVal ) %>% mutate(cohort = factor(cohort))

#newcutoff = round(median(ar$varMAD_median)+sd(ar$varMAD_median)*2, 3)
newcutoff = 0.015

ggplot(ar,aes(sample=varMAD_median)) + stat_qq() + geom_hline(yintercept=newcutoff, linetype='dashed',color='grey') + 
    labs(title='Normal Q-Q plot (all)', x='Theoretical quantiles',y='Sample quantiles') + theme_bw()

ggplot(ar, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
  geom_vline(xintercept = newcutoff, linetype='dashed', color='grey') + labs(x='var(MAD)') +
  theme_bw()

fc = ar %>% filter(round(varMAD_median,3) > newcutoff) %>% select(patient, sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3) %>% arrange(patient, varMAD_median)
```

Using the cutoff `r newcutoff` learned from all of them we would fail `r length(which(resVal$varMAD_median > newcutoff))` samples from the training dataset (Ellie's) and `r length(which(resVal$varMAD_median > newcutoff))` from the initial validation dataset. So while it's less stringent it still fails `r length(which(failedQCVal$varMAD_median > newcutoff))/nrow(resVal)*100`% versus `r length(which(failedQC$varMAD_median > newcutoff))/nrow(res.variance)*100`% of the training dataset.

>>>>

```{r, echo=F, fig.height=10}
res.variance = arrange(res.variance,varMAD_median)
cutoff = round(median(res.variance$varMAD_median)+sd(res.variance$varMAD_median)*3,3)
#cutoff = 0.010

@@

## QC dataset

@@

all.val = all.val %>% rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup
all.val = all.val %>% mutate(
  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
  Sample = paste(`SLX-ID`,sub('-','_',gsub('tp','',`Index Sequence`)),sep='.')
  )

qc.res.val = do.call(bind_rows, lapply(list.files(dir2, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types = 'ccddddl') #%>% set_names('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3', 'n.segs', 'Pass') %>%  
   # mutate(`Patient ID` = as.character(basename(dirname(f))))
})) %>% dplyr::select(patient, samplename, matches('varMad'), -Pass) %>% filter(samplename %in% all.val$Sample)

qc.res.val %>% kable(caption='Residuals across segments per sample') %>% kable_styling('striped')
qc.res.val = arrange(qc.res.val,varMAD_median)

ggplot(qc.res.val,aes(sample=varMAD_median)) + stat_qq() + geom_hline(yintercept=cutoff, linetype='dashed',color='grey') + 
  labs(title='Normal Q-Q plot (all)', x='Theoretical quantiles',y='Sample quantiles') + theme_bw()

ggplot(qc.res.val, aes(varMAD_median)) + geom_histogram(bins=50,color='lightblue') + 
  geom_vline(xintercept = cutoff, linetype='dashed', color='grey') + labs(x='var(median(segments)+1sd(segments))') +
  theme_bw()

failedQCVal = qc.res.val %>% filter(varMAD_median > cutoff) %>% arrange(patient, varMAD_median)

@@

The result from this is that `r nrow(failedQCVal)` samples (`r signif(nrow(failedQCVal)/nrow(qc.res.val)*100,3)`%) would have failed due to high variance across all the segments. A manual check of these agrees with all of them.
